---
title: "Imputation of NHANES Data and Model Selection"
author: "Michael Conlin"
date: "2023-08-01"
output: html_document
---

```{r setup, echo = FALSE, message = FALSE, warning = FALSE}
options(scipen = 1, digits = 4, width = 80, fig.align = "center")
knitr::opts_chunk$set(echo = TRUE)
library(knitr)
```


## Loading Data

Loading the NHANES data for exploration. 

```{r cars}
library(NHANES)
# head(NHANES)
```


### Variable Selection


A look at changing Marijuana to RegularMarijuana and nPregnancies to nBabies:

```{r}
fit_null = lm(BMI ~ Age + Age1stBaby + AlcoholYear + Marijuana + 
    BPSysAve + UrineVol1 + nPregnancies + Poverty + SleepHrsNight + 
    Age:Age1stBaby + Age:AlcoholYear + Age:BPSysAve + Age:UrineVol1 + 
    Age:nPregnancies + Age:SleepHrsNight, 
  data = NHANES)
fit_change = lm(BMI ~ Age + Age1stBaby + AlcoholYear + RegularMarij + 
    BPSysAve + UrineVol1 + nBabies + Poverty + SleepHrsNight + 
    Age:Age1stBaby + Age:AlcoholYear + Age:BPSysAve + Age:UrineVol1 + 
    Age:nBabies + Age:SleepHrsNight, 
  data = NHANES)
summary(fit_null)
summary(fit_change)
```

The model is stronger with the original variable, so I will keep those.

Assessment of LittleInterest vs Depressed

```{r}
table(NHANES$LittleInterest, NHANES$Depressed)
```

These seem to correlate strongly, so no advantage to changing

#### Add in additional possible helpful Variables

I added in Gender, Race1, Education, and MartialStatus.  

```{r}
df_keep = data.frame(predictor = c('SurveyYr', 'Age', 'Age1stBaby', 'AlcoholYear', 'Marijuana', 'SmokeNow', 'HardDrugs', 'BPDiaAve', 'BPSysAve', 'PhysActiveDays', 'TVHrsDay', 'CompHrsDay', 'TotChol', 'Diabetes', 'UrineVol1', 'HealthGen', 'LittleInterest', 'nPregnancies',  'Poverty', 'SleepHrsNight', 'Gender', 'Race1', 'Education', 'MaritalStatus' ))
```


```{r}
nhanes_select = subset(NHANES, select =c(df_keep$predictor, "BMI"))
nrow(NHANES)
ncol(NHANES)
nrow(nhanes_select)
ncol(nhanes_select)
```

We need to know which ones are categorical and turn them into factors

```{r}
nhanes_select$Marijuana = as.factor(nhanes_select$Marijuana)
nhanes_select$SmokeNow = as.factor(nhanes_select$SmokeNow)
nhanes_select$HardDrugs = as.factor(nhanes_select$HardDrugs)
nhanes_select$Diabetes = as.factor(nhanes_select$Diabetes)
nhanes_select$TVHrsDay = as.factor(nhanes_select$TVHrsDay)
nhanes_select$CompHrsDay = as.factor(nhanes_select$CompHrsDay)
nhanes_select$HealthGen = as.factor(nhanes_select$HealthGen)
nhanes_select$LittleInterest = as.factor(nhanes_select$LittleInterest)
nhanes_select$Gender = as.factor(nhanes_select$Gender)
nhanes_select$Race1 = as.factor(nhanes_select$Race1)
nhanes_select$Education = as.factor(nhanes_select$Education)
nhanes_select$MaritalStatus = as.factor(nhanes_select$MaritalStatus)


```


### Multiple Imputation of Data with `mice` Package

```{r eval=FALSE}
if (!require(mice)) {
  install.packages("mice")  
}
```

Here we will perform the imputation. Given the size of the data, this will take a bit of processing time.

```{r}
library(mice)
# remove the rows which have NAs for BMI
nhanes_imp = nhanes_select[!is.na(nhanes_select$BMI), ]
nrow(nhanes_imp)
nrow(nhanes_select)
# perform the multiple imputation (5 datasets)
imp = mice(nhanes_imp, seed = 420, m = 5, print = FALSE)
# Compare the imputed variables (red) and observed (blue)
densityplot(imp)
summary(imp)
```

Now we build and run the stepwise process to select predictors. Notice that our 5 datasets with observed and imputed data are passed to the stepwise function using `with()` which in this case returns a `mira` object from the `mice` package.
```{r}

# build the stepwise workflow
scope <- list(upper = ~ Age +  Age1stBaby +  AlcoholYear +  Marijuana +  SmokeNow +  HardDrugs +  BPDiaAve +  BPSysAve +  PhysActiveDays +  TVHrsDay +  CompHrsDay +  TotChol +  Diabetes +  UrineVol1 +  HealthGen +  LittleInterest +  nPregnancies +   Poverty +  SleepHrsNight +  Gender +  Race1 +  Education +  MaritalStatus + Age:Age1stBaby +  Age:AlcoholYear +  Age:Marijuana +  Age:SmokeNow +  Age:HardDrugs +  Age:BPDiaAve +  Age:BPSysAve +  Age:PhysActiveDays  +  Age:CompHrsDay +  Age:TotChol +  Age:Diabetes +  Age:UrineVol1 +  Age:HealthGen +  Age:LittleInterest +  Age:nPregnancies +   Age:Poverty +  Age:SleepHrsNight +  Age:Gender +  Age:Race1 +  Age:Education +  Age:MaritalStatus,
             lower = ~ 1)
expr <- expression(f1 <- lm(log(BMI) ~ 1),
                  f2 <- step(f1, scope = scope, k = log(nrow(nhanes_select)), trace = 0))
# perform the stepwise selection with each of the 5 imputed datasets
fit <- with(imp, expr)

# count the votes for variables to keep
formulas <- lapply(fit$analyses, formula)
terms <- lapply(formulas, terms)
votes <- unlist(lapply(terms, labels))
table(votes)
```

If we use the criterion of more than half of the datasets resulted in selection of a variable, we end up with:

lm(log(BMI) ~ Age + Age1stBaby + AlcoholYear + BPDiaAve + 
                       BPSysAve + BPSysAve + CompHrsDay + Diabetes + HardDrugs + 
                       HealthGen + nPregnancies + PhysActiveDays  + Race1 +
                       SmokeNow + TVHrsDay + Age:BPDiaAve + Age:BPSysAve +
                       Age:Diabetes + Age:HealthGen + Age:nPregnancies + Age:SmokeNow

Lets see if we can trim out the variables with the lower vote counts (not selected with each imputed dataset)
```{r}
# remove Age:HealthGen
model_with = with(imp, lm(log(BMI) ~ Age + Age1stBaby + AlcoholYear + BPDiaAve + 
                       BPSysAve + BPSysAve + CompHrsDay + Diabetes + HardDrugs + 
                       HealthGen + nPregnancies + PhysActiveDays  + Race1 +
                       SmokeNow + TVHrsDay + Age:BPDiaAve + Age:BPSysAve +
                       Age:Diabetes + Age:HealthGen + Age:nPregnancies + Age:SmokeNow))
model_without = with(imp, lm(log(BMI) ~ Age + Age1stBaby + AlcoholYear + BPDiaAve + 
                       BPSysAve + BPSysAve + CompHrsDay + Diabetes + HardDrugs + 
                       HealthGen + nPregnancies + PhysActiveDays  + Race1 +
                       SmokeNow + TVHrsDay + Age:BPDiaAve + Age:BPSysAve +
                       Age:Diabetes + Age:nPregnancies + Age:SmokeNow))
anova(model_without, model_with)
```

The p-value suggests that Age:HealthGen is not worth keeping.

```{r}
# remove Age:nPregnancies
model_with = with(imp, lm(log(BMI) ~ Age + Age1stBaby + AlcoholYear + BPDiaAve + 
                       BPSysAve + BPSysAve + CompHrsDay + Diabetes + HardDrugs + 
                       HealthGen + nPregnancies + PhysActiveDays  + Race1 +
                       SmokeNow + TVHrsDay + Age:BPDiaAve + Age:BPSysAve +
                       Age:Diabetes + Age:nPregnancies + Age:SmokeNow))
model_without = with(imp, lm(log(BMI) ~ Age + Age1stBaby + AlcoholYear + BPDiaAve + 
                       BPSysAve + BPSysAve + CompHrsDay + Diabetes + HardDrugs + 
                       HealthGen + nPregnancies + PhysActiveDays  + Race1 +
                       SmokeNow + TVHrsDay + Age:BPDiaAve + Age:BPSysAve +
                       Age:Diabetes + Age:SmokeNow))
anova(model_without, model_with)
```

With a p-value of less than 0.01, lets keep Age:nPregnancies

```{r}
# remove HardDrugs
model_with = with(imp, lm(log(BMI) ~ Age + Age1stBaby + AlcoholYear + BPDiaAve + 
                       BPSysAve + BPSysAve + CompHrsDay + Diabetes + HardDrugs + 
                       HealthGen + nPregnancies + PhysActiveDays  + Race1 +
                       SmokeNow + TVHrsDay + Age:BPDiaAve + Age:BPSysAve +
                       Age:Diabetes + Age:HealthGen + Age:nPregnancies + Age:SmokeNow))
model_without = with(imp, lm(log(BMI) ~ Age + Age1stBaby + AlcoholYear + BPDiaAve + 
                       BPSysAve + BPSysAve + CompHrsDay + Diabetes + 
                       HealthGen + nPregnancies + PhysActiveDays  + Race1 +
                       SmokeNow + TVHrsDay + Age:BPDiaAve + Age:BPSysAve +
                       Age:Diabetes + Age:HealthGen+ Age:nPregnancies  + Age:SmokeNow))
anova(model_without, model_with)
```

It also seems that HardDrugs is not needed in our model.

```{r}
# remove nPregnancies
model_with =with(imp, lm(log(BMI) ~ Age + Age1stBaby + AlcoholYear + BPDiaAve + 
                       BPSysAve + BPSysAve + CompHrsDay + Diabetes + 
                       HealthGen + nPregnancies + PhysActiveDays  + Race1 +
                       SmokeNow + TVHrsDay + Age:BPDiaAve + Age:BPSysAve +
                       Age:Diabetes + Age:HealthGen + Age:SmokeNow))
model_without = with(imp, lm(log(BMI) ~ Age + Age1stBaby + AlcoholYear + BPDiaAve + 
                       BPSysAve + BPSysAve + CompHrsDay + Diabetes + 
                       HealthGen + PhysActiveDays  + Race1 +
                       SmokeNow + TVHrsDay + Age:BPDiaAve + Age:BPSysAve +
                       Age:Diabetes + Age:HealthGen + Age:SmokeNow))
anova(model_without, model_with)
```

With a p-value greater than 0.01, we can safely remove nPregnacies.

### Summary table of 5 models fitted so far

```{r}
library(lmtest)
calc_loocv_rmse = function(model) {
  sqrt(mean((resid(model) / (1 - hatvalues(model))) ^ 2))
}

fit_summary <- data.frame(adj_r_squared = rep(0, 5), loocv_rmse = rep(0, 5), bptest_p = rep(1,5))
for (i in seq(1,5)) {
  fit_summary$adj_r_squared[i] = summary(fit$analyses[[i]])$adj
  fit_summary$loocv_rmse[i] = calc_loocv_rmse(fit$analyses[[i]]) 
  fit_summary$bptest_p[i] = unname(bptest(fit$analyses[[i]])$p.value)
  kable(data.frame(summary(fit$analyses[[i]])$coefficients[, "Estimate"]))
}
kable(fit_summary)
```



```{r}
kable(data.frame(summary(fit$analyses[[1]])$coefficients))
kable(data.frame(summary(fit$analyses[[2]])$coefficients))
kable(data.frame(summary(fit$analyses[[3]])$coefficients))
kable(data.frame(summary(fit$analyses[[4]])$coefficients))
kable(data.frame(summary(fit$analyses[[5]])$coefficients))
```

Functions for diagnostics and model assesments

```{r}

fit_summary <- data.frame(adj_r_squared = rep(0, 5), loocv_rmse = rep(0, 5), bptest_p = rep(1,5), ad_test = rep(0,5))
for (i in seq(1,5)) {
  fit_summary$adj_r_squared[i] = summary(fit_add$analyses[[i]])$adj
  fit_summary$loocv_rmse[i] = calc_loocv_rmse(fit_add$analyses[[i]]) 
  fit_summary$bptest_p[i] = unname(bptest(fit_add$analyses[[i]])$p.value)
  ad.test(residuals(fit_add$analyses[[i]]))$p.value # Anderson-Darling test for normal distribution in larger sample sizes (> 5000)
}
# model diagnostics
model_diagnostics = function(fit){
  fit_summary <- data.frame(bptest_p = rep(0,5), ad_test = rep(0,5))
  for (i in 1:5){
    fit_summary$bptest_p[i] = unname(bptest(fit$analyses[[i]])$p.value)
    ad.test(residuals(fit$analyses[[i]]))$p.value 
  }
  kable(fit_summary, col.names = c("BP Test", "AD Test"))
}
model_diagnostics(fit_add)

# cooks distance to check for influential observations
cooks_function = function(fit){
  cook_thresh = 4 * cooks.distance(fit$analyses[[1]]) / nrow(imp)
  mean(cooks.distance(fit$analyses[[1]]) > cook_thresh)
}
 
cooks_function(fit_add)

# Function to calculate the LOOCVRMSE
calc_loocv_rmse = function(model) {
  sqrt(mean((resid(model) / (1 - hatvalues(model))) ^ 2))
}

# model assessments
model_assess = function(fit){
  fit_summary <- data.frame(adj_r_squared = rep(0,5), loocv_rmse = rep(0,5))
  for (i in 1:5){
    fit_summary$adj_r_squared[i] = summary(fit$analyses[[i]])$adj
    fit_summary$loocv_rmse[i] = calc_loocv_rmse(fit$analyses[[i]]) 
  }
  kable(fit_summary, col.names = c("Adj. R-Squared", "LOOCV-RMSE"))
}

model_assess(fit_add)
```


### Consider log transformations of numerical measurements

BMI

```{r}
par(mfrow=c(1,2))
hist(NHANES$BMI,
     main = "Histogram - BMI",
     xlab = "BMI")
hist(log(NHANES$BMI),
     main = "Histogram - Log(BMI)",
     xlab = "Log(BMI)")
```

BPSysAve

```{r}
par(mfrow=c(1,2))
hist(NHANES$BPSysAve,
     main = "Histogram - BPSysAve",
     xlab = "BPSysAve")
hist(log(NHANES$BPSysAve),
     main = "Histogram - Log(BPSysAve)",
     xlab = "Log(BPSysAve)")
```
BPDiasAve

```{r}
par(mfrow=c(1,2))
hist(NHANES$BPDiaAve,
     main = "Histogram - BPDiaAve",
     xlab = "BPDiaAve")
hist((NHANES$BPDiaAve)^2,
     main = "Histogram - BPDiaAve^2",
     xlab = "BPDiaAve^2")
```

TotChol

```{r}
par(mfrow=c(1,2))
hist(NHANES$TotChol,
     main = "Histogram - TotChol",
     xlab = "TotChol")
hist(log(NHANES$TotChol),
     main = "Histogram - TotChol",
     xlab = "TotChol")
```
```{r}
par(mfrow=c(1,2))
hist(NHANES$SleepHrsNight,
     main = "Histogram - SleepHrsNight",
     xlab = "SleepHrsNight")
hist(log(NHANES$SleepHrsNight),
     main = "Histogram - SleepHrsNight",
     xlab = "SleepHrsNight")
```

```{r}
hist((NHANES$Poverty))
```

#### Conclusions

Based on the above findings, we may want to log transform BPSysAve, and TotChol, and consider a square transform of BPDiaAve. I think SleepHrsNight is best left alone, and I do not know what to make of the Poverty levels.


#### Compare smaller model 

```{r}
fit_add_aic_with = with(imp, lm(log(BMI) ~ Age +  AlcoholYear + SmokeNow +
  BPDiaAve + BPSysAve + TVHrsDay + CompHrsDay +
  TotChol +  Diabetes +  UrineVol1 +  HealthGen +  LittleInterest +
  SleepHrsNight +  Race1  +  MaritalStatus))
# removing UrineVol1 
fit_add_aic_without =with(imp, lm(log(BMI) ~ Age +  AlcoholYear + SmokeNow +
  BPDiaAve + BPSysAve + TVHrsDay + CompHrsDay +
  TotChol +  Diabetes +  HealthGen +  LittleInterest +
  SleepHrsNight +  Race1  +  MaritalStatus))
anova(fit_add_aic_without, fit_add_aic_with)
```

```{r}
fit_add_aic_with = with(imp, lm(log(BMI) ~ Age +  AlcoholYear + SmokeNow +
  BPDiaAve + BPSysAve + TVHrsDay + CompHrsDay +
  TotChol +  Diabetes +  HealthGen +  LittleInterest +
  SleepHrsNight +  Race1  +  MaritalStatus))
# removing TocChol
fit_add_aic_without =with(imp, lm(log(BMI) ~ Age +  AlcoholYear + SmokeNow +
  BPDiaAve + BPSysAve + TVHrsDay + CompHrsDay +
  Diabetes +  HealthGen +  LittleInterest +
  SleepHrsNight +  Race1  +  MaritalStatus))
anova(fit_add_aic_without, fit_add_aic_with)
```

#### Try step back with added interaction terms


```{r}
# build the stepwise workflow using our fit_add_aic 
# with interaction terms added as a starting point
scope <- list(upper = ~ Age +  AlcoholYear + SmokeNow +
  BPDiaAve + BPSysAve + TVHrsDay + CompHrsDay +
  TotChol +  Diabetes +  UrineVol1 +  HealthGen +  LittleInterest +
  SleepHrsNight +  Race1  +  MaritalStatus +  Age:AlcoholYear +  Age:SmokeNow +  Age:BPDiaAve +  Age:BPSysAve  + Age:TVHrsDay + Age:CompHrsDay +  Age:TotChol +  Age:Diabetes +  Age:UrineVol1 +  Age:HealthGen +  Age:LittleInterest +  Age:SleepHrsNight +  Age:Race1 +  Age:MaritalStatus,
             lower = ~ 1)
expr <- expression(f1 <- lm(log(BMI) ~ 1),
                  f2 <- step(f1, scope = scope, trace = 0))
# perform the stepwise selection with each of the 5 imputed datasets
fit <- with(imp, expr)

# count the votes for variables to keep
formulas <- lapply(fit$analyses, formula)
terms <- lapply(formulas, terms)
votes <- unlist(lapply(terms, labels))
table(votes)

```

Build the model with all the above predictors with > 3 votes

```{r}
fit_int_aic = with(imp, lm(log(BMI) ~ Age +  AlcoholYear + SmokeNow +
  BPDiaAve + BPSysAve + TVHrsDay + CompHrsDay +
  Diabetes +  HealthGen +  LittleInterest +
  SleepHrsNight +  Race1  +  MaritalStatus +  Age:SmokeNow +  Age:BPDiaAve +  Age:BPSysAve  + Age:TVHrsDay  +  Age:Diabetes +   Age:HealthGen +  Age:Race1 +  Age:MaritalStatus))
summary(fit_int_aic$analyses[[1]])
```

Consider taking out `Age:TVHrsDay`

```{r}
fit_int_aic_with = with(imp, lm(log(BMI) ~  Age +  AlcoholYear + SmokeNow +
  BPDiaAve + BPSysAve + TVHrsDay + CompHrsDay +
  Diabetes +  HealthGen +  LittleInterest +
  SleepHrsNight +  Race1  +  MaritalStatus +  Age:SmokeNow +  Age:BPDiaAve +  Age:BPSysAve  + Age:TVHrsDay  +  Age:Diabetes +   Age:HealthGen +  Age:Race1 +  Age:MaritalStatus))
# removing Age:TVHrsDay
fit_int_aic_without =with(imp, lm(log(BMI) ~  Age +  AlcoholYear + SmokeNow +
  BPDiaAve + BPSysAve + TVHrsDay + CompHrsDay +
  Diabetes +  HealthGen +  LittleInterest +
  SleepHrsNight +  Race1  +  MaritalStatus +  Age:SmokeNow +  Age:BPDiaAve +  Age:BPSysAve  +  Age:Diabetes +   Age:HealthGen +  Age:Race1 +  Age:MaritalStatus))
anova(fit_int_aic_without, fit_int_aic_with)
```


log transformation of BP measures

```{r}
fit_int_aic_trns = with(imp, lm(log(BMI) ~ Age +  AlcoholYear + SmokeNow +
  I(BPDiaAve^2) + log1p(BPSysAve) + TVHrsDay + CompHrsDay +
  Diabetes +  HealthGen +  LittleInterest +
  SleepHrsNight +  Race1  +  MaritalStatus +  Age:SmokeNow +  Age:I(BPDiaAve^2) +  Age:log1p(BPSysAve)  + Age:TVHrsDay + Age:CompHrsDay  +  Age:Diabetes +  Age:HealthGen +  Age:Race1 +  Age:MaritalStatus))
summary(fit_int_aic_trns$analyses[[1]])
```

```{r}

sum(cooks.distance(fit_int_aic_trns$analyses[[3]]) > 4 / length(cooks.distance(fit_int_aic_trns$analyses[[3]])))

imp_df_1 = imp_df[imp_df$.imp == 1, ]

test = lm(log(BMI) ~ Age +  AlcoholYear + SmokeNow +
  I(BPDiaAve^2) + log1p(BPSysAve) + TVHrsDay + CompHrsDay +
  Diabetes +  HealthGen +  LittleInterest +
  SleepHrsNight +  Race1  +  MaritalStatus +  Age:SmokeNow +  Age:I(BPDiaAve^2) +  Age:log1p(BPSysAve)  + Age:TVHrsDay + Age:CompHrsDay  +  Age:Diabetes +  Age:HealthGen +  Age:Race1 +  Age:MaritalStatus, data = imp_df_1)
summary(test)
```

